Statistical Guarantees for Local Spectral Clustering on Random Neighborhood Graphs

### 2021-01-24 jmlr/revision/local_spectral_clustering_jmlr_revision_2.tex
Addressed (nearly) all of Ryan's comments on the revised manuscript.

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

@@

\subsection{PPR clustering, and worst-case guarantees}
We now review the PPR clustering algorithm, and some worst-case guarantees on the quality of the resulting cluster. Let $G = (V,E)$ be an undirected, unweighted, and connected graph. We denote by $A \in \Reals^{n \times n}$ the adjacency
matrix of $G$, with entries $A_{uv} = 1$ if $(u,v) \in E$ and $0$ otherwise.  We
also denote by $D$ the diagonal degree matrix, with entries $D_{uu} :=
\sum_{v \in V} A_{uv}$, and by $I$ the $n \times n$ identity matrix. The PPR vector $p_v = p(v,\alpha;G)$ is defined with respect to a given seed node $v \in V$ and a teleportation parameter $\alpha \in [0,1]$, as the solution of the following linear system:
\begin{equation}
\label{eqn:ppr_vector}
p_v = \alpha e_{v} + (1 - \alpha) p_v W,
\end{equation}
where $W = (I + D^{-1}A)/2$ is the lazy random walk matrix over
$G$ and $e_{v}$ is the indicator vector for node $v$ (that has a 1 in
position $v$ and 0 elsewhere).

In practice, exactly solving the system of equations~\eqref{eqn:ppr_vector} to compute the PPR vector may be too computationally expensive. To address this limitation, \citet{andersen2006} introduced the \emph{$\varepsilon$-approximate} PPR vector (aPPR), which we will denote by \smash{$p^{(\varepsilon)}$}. We refer the curious reader to \citet{andersen2006} for a formal algorithmic definition of the aPPR vector, and limit ourselves to highlighting a few salient points: the aPPR vector can be computed in order $\mathcal{O}(1/(\varepsilon\alpha))$ time, while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn: appr_error}
\textrm{for all $u \in V$}, \quad p_v(u) - \varepsilon D_{uu}\leq
p_v^{(\varepsilon)}(u) \leq p_v(u).  
\end{equation}

@@

As mentioned, to date most analysis of local clustering has focused on worst-case guarantees, defined with respect to functionals of an a priori fixed graph $G = (V,E)$.  For instance, \cite{andersen2006} analyze the normalized cut of the cluster estimate $\wh{C}$ output by PPR, showing that when PPR is appropriately seeded within a candidate cluster $C \subseteq V$, the normalized cut of $\wh{C}$ is upper bounded $\Phi(\wh{C};G) \lesssim \sqrt{\Phi(C;G)}$. \cite{zhu2013} build on this: they introduce a second functional, the \emph{conductance} $\Psi(G)$, defined as 
\begin{equation}
\label{eqn:conductance}
\Psi(G) := \min_{S \subseteq V} \Phi(S;G),
\end{equation}
and show that if $\Phi(C;G)$ is much smaller than $\Psi(G[C])^2$---where $G[C] = (C,E \cap (C \times C))$ is the subgraph of $G$ induced by $C$--- then (in addition to having a small normalized cut) the cluster estimate $\wh{C}$ approximately recovers $C$. Our own analysis builds on that of~\cite{zhu2013}, and we give a more detailed summary of their results in Section~\ref{sec:ub_symmetric_set_difference}.
For now, we merely reiterate that the conclusions of \citep{andersen2006,zhu2013} cannot be straightforwardly applied to our statistical setting, where the input data are random samples $\{x_1,\ldots,x_n\}$ drawn from a distribution $\mathbb{P}$, the graph $G$ is a random geometric graph formed by the user, and the candidate cluster is a set $\mc{C} \subset \Rd$.\footnote{Throughout, we use calligraphic notation to refer to subsets of $\Rd$.}

\subsection{PPR on a neighborhood graph}
We now formally describe the statistical setting in which we operate, as well as the method we will study: PPR on a neighborhood graph. Let $X = \{x_1,\ldots, x_n\}$ be samples drawn i.i.d.\ from a distribution
$\Pbb$ on $\Rd$. We will assume throughout that $\Pbb$ has a density $f$ with respect to the Lebesgue measure $\nu$ on $\Rd$ . For a radius $r > 0$, we define
$G_{n,r}=(V,E)$ to be the \emph{$r$-neighborhood graph} of $X$, an
unweighted, undirected graph with vertices $V=X$, and an edge $(x_i,x_j) \in
E$ if and only if $\|x_i - x_j\| \leq r$, where $\|\cdot\|$ is the
Euclidean norm. Once the neighborhood graph $G_{n,r}$ is formed, the PPR vector $p_v$ is then computed over $G_{n,r}$, with a resulting cluster estimate $\wh{C} \subseteq X$. The precise PPR algorithm we analyze is summarized in Algorithm~\ref{alg:ppr}.

@@

\subsection{Cluster estimation}
We need a metric to assess the accuracy with which $\wh{C}$ estimates the candidate cluster $\mc{C}$. One commonly used metric is the misclassification error, i.e. the size of the symmetric set difference
between $\wh{C}$ and the empirical cluster $\mc{C}[X] = \mc{C} \cap X$ \citep{korostelev1993,polonik1995,rigollet2009}. We
will consider a related metric, the volume of the symmetric set difference,
which weights misclassified points according to their degree in $G_{n,r}$. To keep things simple, for a given set $S \subseteq X$ we write $\vol_{n,r}(S) := \vol(S;G_{n,r})$. 
\begin{definition}
	\label{def:volume_symmetric_set_difference}
	For an estimator \smash{$\wh{C} \subseteq X$} and a set 
	$\mathcal{C} \subseteq \Rd$, their symmetric set difference is 
	\begin{equation*}
	\wh{C} \vartriangle \mathcal{C}[X] :=
	\bigl(\wh{C} \setminus \mathcal{C}[X]\bigr) \cup
	\bigl(\mathcal{C}[X] \setminus \wh{C}\bigr).
	\end{equation*}
	Furthermore, we denote the volume of the symmetric set difference by 
	$$
	\Delta(\wh{C}, \mathcal{C}[X]) := \vol_{n,r}(\wh{C} \vartriangle \mathcal{C}[X]). 
	$$
\end{definition}

@@

It is quite natural that $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$ should help quantify the role geometry plays in local spectral clustering. Indeed, by construction these functionals are quite obviously the population-level analogues of the empirical quantities $\Phi_{n,r}(\mc{C}[X]) := \Phi(\mc{C}[X];G_{n,r})$ and $\Psi_{n,r}(\mc{C}[X]) = \Psi(G_{n,r}\bigl[\mc{C}[X]\bigr])$, and as we have already mentioned, these empirical quantities in turn suffice to upper bound the volume of the symmetric set difference. For this reason, similar population level functionals are used by \citep{shi2009,schiebinger2015,garciatrillos19} in the analysis of \emph{global} spectral clustering in a statistical context. We will comment more on the relationship between these works and our own results in Section~\ref{subsec:related_work}. \\

@@

\subsection{Summary of results}

A summary of our results (and an outline for this paper) is as follows.
\begin{itemize}
	\item In Section~\ref{sec:ub_symmetric_set_difference}, we show that under certain conditions the sample normalized cut $\Phi_{n,r}(\mc{C}[X])$ and conductance $\Psi_{n,r}(\mc{C}[X])$ are close to their population level counterparts $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$. As a consequence, we find that $\Delta(\wh{C},\mc{C}[X])$ is small whenever $\Phi_{\Pbb,r}(\mc{C})$ is small relative to $\Psi_{\Pbb,r}(\mc{C})^2$, giving the first population-level guarantees for local clustering in a nonparametric statistical context.
\end{itemize}
In Section~\ref{sec:ppr_density_cluster}, we focus on the special case where the candidate cluster $\mc{C} = \mc{C}_{\lambda}$ is a \emph{$\lambda$-density cluster}---that is, a connected component of the upper level set $\{x: f(x) \geq \lambda\}$---an object of central interest in the statistical clustering literature, dating back to the work of \citet{hartigan1981}.
\begin{itemize}
	\item In Section~\ref{subsec:recovery_well-conditioned_density_clusters}, we derive specific upper bounds on the normalized cut $\Phi_{\Pbb,r}(\mc{C}_{\lambda})$ and conductance $\Psi_{\Pbb,r}(\mc{C}_{\lambda})$. These upper bounds depend on $\lambda$ as well as some other natural parameters, and immediately result in an upper bound on $\Delta(\wh{C},\mc{C}[X])$ in the density cluster setting (Corollary~\ref{cor:density_cluster_volume_ssd_ub}).
	\item In Section~\ref{subsec:lower_bound}, we prove a negative result: we give a hard distribution $\Pbb$ with corresponding density cluster $\mc{C}_{\lambda}$ for which the symmetric set difference $\Delta(\wh{C},\mc{C}_{\lambda}[X])$ is provably large.
\end{itemize}
In Section~\ref{sec:experiments} we empirically investigate some of our conclusions, before ending with some discussion in Section~\ref{sec:discussion}.

\subsection{Related Work}
\label{subsec:related_work}
First, however, we conclude this Section by summarizing some related work (in addition to the background already given above). In the stochastic block model (SBM), arguably one of the simplest models of network formation, edges between nodes independently occur with probability based on a latent community membership. In the SBM, the ability of spectral algorithms to perform clustering---or community detection---is well-understood, dating 
back to \citet{mcsherry2001} who gives conditions under which the entire
community structure can be recovered. In more recent work, \citet{rohe2011}
upper bound the fraction of nodes misclassified by a spectral algorithm for the
high-dimensional (large number of blocks) SBM, and \citet{lei2015} extend these
results to the sparse (low average degree) regime. Relatedly,
\citet{clauset08,balakrishnan2011,li2018}, analyze the misclassification rate
when the block model exhibits some hierarchical structure. The framework we
consider, in which nodes correspond to data points sampled from an underlying 
density, and edges between nodes are formed based on geometric proximity, is
quite different than the SBM, and therefore so is our analysis.

@@

\section{Upper bound on the symmetric set difference metric}
\label{sec:ub_symmetric_set_difference}

In the main result (Theorem~\ref{thm:volume_ssd_ub}) of this section, we give a high probability upper bound on $\Delta(\wh{C}, \mc{C}[X])$, in terms of the population normalized cut $\Phi_{\Pbb,r}(\mc{C})$ and conductance $\Psi_{\Pbb,r}(\mc{C})$. We build to this Theorem slowly, giving new structural results in two distinct directions. First, we build on some previous work (mentioned in the Introduction) to relate $\Delta(\wh{C}, \mc{C}[X])$ to the sample normalized cut $\Phi_{n,r}(\mc{C}[X])$ and conductance $\Psi_{n,r}(\mc{C}[X])$. Second, we argue that when $n$ is large, $\Phi_{n,r}(\mc{C}[X])$ and $\Psi_{n,r}(\mc{C}[X])$ can be suitably bounded by their population-level analogues $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$.

\subsection{PPR cluster recovery: the fixed graph case.}
\label{subsec:ppr_cluster_recovery_fixed_graph}
When PPR is run on a fixed graph $G = (V,E)$ with the goal of recovering a candidate cluster $C \subset V$, \cite{zhu2013} provide the sharpest known bounds on the volume of the symmetric set difference between the cluster estimate $\wh{C}$ and candidate cluster $C$. Since these results will play a major part in our analysis, in Lemma~\ref{lem:zhu} we restate them for the convenience of the reader.\footnote{Lemma~\ref{lem:zhu} improves on Lemma~3.4 of~\cite{zhu2013} by some constant factors, and for completeness we prove Lemma~\ref{lem:zhu} in our appendix. Nevertheless, to be clear the essential idea of Lemma~\ref{lem:zhu} is no different than that of~\cite{zhu2013}, and we do not claim any novelty.}

@@

\paragraph{Improved bounds on mixing time.} Having reviewed the conclusions of~\cite{zhu2013}, we return now to our own setting, where the data is not a fixed graph $G$ but instead random samples $\{x_1,\ldots,x_n\}$, and our goal is to recover a candidate cluster $\mc{C} \subset \Rd$. Ideally, we would like to apply~\eqref{eqn:zhu_ub2} with $C = \mc{C}[X]$ and $G = G_{n,r}$, replace $\Phi_{n,r}(\mc{C}[X])$ and $\Psi_{n,r}(\mc{C}[X])$ by $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$ inside~\eqref{eqn:zhu_ub2}, and thereby obtain an upper bound on $\Delta(\wh{C};\mc{C}[X])$ that depends only on $\Pbb$ and $\mc{C}$. Unfortunately, however, there is a catch: when the graph $G = G_{n,r}$ and the candidate cluster $C = \mc{C}[X]$, as $n \to \infty$ the sample normalized cut $\Phi_{n,r}(\mc{C}[X])$ and conductance $\Psi_{n,r}(\mc{C}[X])$ each converge to their population-level analogues, but $\pi_{\min}\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr) \asymp 1/n$. Therefore the right hand side of~\eqref{eqn:zhu_ub2} grows at a $\log n$ rate, rendering~\eqref{eqn:zhu_ub2} a vacuous upper bound when the number of samples is sufficiently large.

@@

To establish these bounds, we impose the following regularity conditions on $\wt{\Pbb}$ and $\mc{C}$.
\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:domain} 
	The candidate cluster $\mc{C} \subseteq \Rd$ for $d \geq 2$, is a bounded, connected, open set with Lipschitz boundary. 
	\item 
	\label{asmp:bounded_density} 
	The distribution $\wt{\Pbb}$ has a density $\wt{f}: \mc{C} \to (0,\infty)$ such that there exist $f_{\min} \leq 1 \leq f_{\max}$ for which
	\begin{equation*}
	(\forall x \in \mc{C})~~ f_{\min} \leq \wt{f}(x) \leq f_{\max}.
	\end{equation*}
\end{enumerate}
In what follows, we use $b_1,b_2,\ldots$ and $B_1,B_2,\ldots$ to refer to constants that may depend on $\Pbb$ and $\mc{C}$, but do not depend on $n$ or $\delta$. We explicitly keep track of all constants in our proofs.
\begin{proposition}
	\label{prop:sample_to_population_1}
	Fix $\delta \in (0,1/3)$. Suppose $\mc{C}$ and $\wt{\Pbb}$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_density}. There exist positive constants $b_1$-$b_3$ such that each of the following statements hold.
	\begin{itemize}
		\item With probability at least $1 - 3\exp\{-b_1\delta^2n\}$,
		\begin{equation}
		\label{eqn:sample_to_population_normalized_cut}
		\Phi_{n,r}(\mc{C}[X]) \leq (1 + 3\delta) \Phi_{\Pbb,r}(\mc{C}).
		\end{equation}
		\item With probability at least $1 - 2\exp\{-b_2\delta^2n\} - n\exp\{-b_3\delta^2n\}$,
		\begin{equation}
		\label{eqn:sample_to_population_local_spread}
		s_{n,r}(\mc{C}[X]) \geq (1 - 3\delta) s_{\Pbb,r}(\mc{C}).
		\end{equation}
	\end{itemize}
\end{proposition}
Let $p_d := 3/4$ if $d = 2$, and else $p_d := 1/d$ for $d \geq 3$. 
\begin{proposition}
	\label{prop:sample_to_population_2}
	Fix $\delta \in (0,1/2)$. Suppose $\mc{C}$ and $\wt{\Pbb}$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_density}. Then there exist positive constants $B_1$ and $B_2$, such that for any $n \in \mathbb{N}$ for which
	\begin{equation}
	\label{eqn:sample_to_population_conductance_sample_complexity}
	\frac{\log(n)^{p_d}}{n^{1/d}} \leq \delta \cdot B_1,
	\end{equation}
	the following inequality holds with probability at least $1 - B_2/n - (n + 2)\exp\{-b_2n/\sqrt{2}\}$:
	\begin{equation}
	\label{eqn:sample_to_population_conductance}
	\Psi_{n,r}(\mc{C}[X]) \geq (1 - 2\delta) \Psi_{\Pbb,r}(\mc{C}).
	\end{equation}
\end{proposition}

@@

\paragraph{Well-initialized algorithm.} As is typical in the local clustering literature, our algorithmic results will be stated with respect to specific ranges of each of the user-specified
parameters. In particular, for $\delta \in (0,1/4)$ and a candidate cluster $\mc{C} \in \Rd$, we require that some of the tuning parameters of Algorithm~\ref{alg:ppr} be chosen within specific ranges,

>>>>

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ubb}{\mathbb{U}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

@@

\subsection{PPR clustering}
\label{subsec:ppr_clustering}
We now review the PPR clustering algorithm, and some worst-case guarantees on the quality of the resulting cluster. Let $G = (V,E)$ be an undirected, unweighted, and connected graph. We denote by $A \in \Reals^{n \times n}$ the adjacency
matrix of $G$, with entries $A_{uv} = 1$ if $(u,v) \in E$ and $0$ otherwise.  We
also denote by $D$ the diagonal degree matrix, with entries $D_{uu} :=
\sum_{v \in V} A_{uv}$, and by $I$ the $n \times n$ identity matrix. The PPR vector $p_v = p(v,\alpha;G)$ is defined with respect to a given seed node $v \in V$ and a teleportation parameter $\alpha \in [0,1]$, as the solution of the following linear system:
\begin{equation}
\label{eqn:ppr_vector}
p_v = \alpha e_{v} + (1 - \alpha) p_v W,
\end{equation}
where $W = (I + D^{-1}A)/2$ is the lazy random walk matrix over
$G$ and $e_{v}$ is the indicator vector for node $v$ (that has a 1 in
position $v$ and 0 elsewhere).

In practice, exactly solving the system of equations~\eqref{eqn:ppr_vector} to compute the PPR vector may be too computationally expensive. To address this limitation, \citet{andersen2006} introduced the \emph{$\varepsilon$-approximate} PPR vector (aPPR), which we will denote by \smash{$p^{(\varepsilon)}$}. We refer the curious reader to \citet{andersen2006} for a formal algorithmic definition of the aPPR vector, and limit ourselves to highlighting a few salient points: the aPPR vector can be computed in order $\mathcal{O}(1/(\varepsilon\alpha))$ time, while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn:appr_error}
\textrm{for all $u \in V$}, \quad p_v(u) - \varepsilon D_{uu}\leq
p_v^{(\varepsilon)}(u) \leq p_v(u).  
\end{equation}

@@

\subsection{Worst-case guarantees for PPR clustering}
As mentioned, to date most analysis of local clustering has focused on worst-case guarantees, defined with respect to functionals of an a priori fixed graph $G = (V,E)$.  For instance, \cite{andersen2006} analyze the normalized cut of the cluster estimate $\wh{C}$ output by PPR, showing that when PPR is appropriately seeded within a candidate cluster $C \subseteq V$, the normalized cut of $\wh{C}$ is upper bounded $\Phi(\wh{C};G) \lesssim \sqrt{\Phi(C;G)}$.\footnote{We say $a \lesssim b$ when $a \leq $} \cite{zhu2013} build on this: they introduce a second functional, the \emph{conductance} $\Psi(G)$, defined as 
\begin{equation}
\label{eqn:conductance}
\Psi(G) := \min_{S \subseteq V} \Phi(S;G),
\end{equation}
and show that if $\Phi(C;G)$ is much smaller than $\Psi(G[C])^2$---where $G[C] = (C,E \cap (C \times C))$ is the subgraph of $G$ induced by $C$--- then (in addition to having a small normalized cut) the cluster estimate $\wh{C}$ approximately recovers $C$. Our own analysis builds on that of~\cite{zhu2013}, and we give a more detailed summary of their results in Section~\ref{sec:ub_symmetric_set_difference}.
For now, we merely reiterate that the conclusions of \citep{andersen2006,zhu2013} cannot be straightforwardly applied to our statistical setting, where the input data are random samples $\{x_1,\ldots,x_n\}$ drawn from a distribution $\mathbb{P}$, the graph $G$ is a random neighborhood graph formed by the user, and the candidate cluster is a set $\mc{C} \subset \Rd$.\footnote{Throughout, we use calligraphic notation to refer to subsets of $\Rd$.}

\subsection{PPR on a neighborhood graph}
\label{subsec:ppr_neighborhood_graph}
We now formally describe the statistical setting in which we operate, as well as the method we will study: PPR on a neighborhood graph. Let $X = \{x_1,\ldots, x_n\}$ be samples drawn i.i.d.\ from a distribution
$\Pbb$ on $\Rd$. We will assume throughout that $\Pbb$ has a density $f$ with respect to the Lebesgue measure $\nu$ on $\Rd$ . For a radius $r > 0$, we define
$G_{n,r}=(V,E)$ to be the \emph{$r$-neighborhood graph} of $X$, an
unweighted, undirected graph with vertices $V=X$, and an edge $(x_i,x_j) \in
E$ if and only if $\|x_i - x_j\| \leq r$, where $\|\cdot\|$ is the
Euclidean norm. Once the neighborhood graph $G_{n,r}$ is formed, the PPR vector $p_v$ is then computed over $G_{n,r}$, with a resulting cluster estimate $\wh{C} \subseteq X$. The precise PPR algorithm we analyze is summarized in Algorithm~\ref{alg:ppr}.

@@

\subsection{Cluster accuracy}
We need a metric to assess the accuracy with which $\wh{C}$ estimates the candidate cluster $\mc{C}$. One commonly used metric is the misclassification error, i.e. the size of the symmetric set difference
between $\wh{C}$ and the empirical cluster $\mc{C}[X] = \mc{C} \cap X$ \citep{korostelev1993,polonik1995,rigollet2009}. We
will consider a related metric, the volume of the symmetric set difference,
which weights misclassified points according to their degree in $G_{n,r}$. To keep things simple, for a given set $S \subseteq X$ we write $\vol_{n,r}(S) := \vol(S;G_{n,r})$. 
\begin{definition}
	\label{def:volume_symmetric_set_difference}
	For an estimator \smash{$\wh{C} \subseteq X$} and a set 
	$\mathcal{C} \subseteq \Rd$, their symmetric set difference is 
	\begin{equation*}
	\wh{C} \vartriangle \mathcal{C}[X] :=
	\bigl(\wh{C} \setminus \mathcal{C}[X]\bigr) \cup
	\bigl(\mathcal{C}[X] \setminus \wh{C}\bigr).
	\end{equation*}
	Furthermore, we denote the volume of the symmetric set difference by 
	$$
	\Delta(\wh{C}, \mathcal{C}[X]) := \vol_{n,r}(\wh{C} \vartriangle \mathcal{C}[X]). 
	$$
\end{definition}

@@

It is quite natural that $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$ should help quantify the role geometry plays in local spectral clustering. Indeed, by construction these functionals are quite obviously the population-level analogues of the empirical quantities $\Phi_{n,r}(\mc{C}[X]) := \Phi(\mc{C}[X];G_{n,r})$ and $\Psi_{n,r}(\mc{C}[X]) := \Psi(G_{n,r}\bigl[\mc{C}[X]\bigr])$, and as we have already mentioned, these empirical quantities in turn suffice to upper bound the volume of the symmetric set difference. For this reason, similar population level functionals are used by \citep{shi2009,schiebinger2015,garciatrillos19} in the analysis of \emph{global} spectral clustering in a statistical context. We will comment more on the relationship between these works and our own results in Section~\ref{subsec:related_work}. \\

@@

\subsection{Summary of results}
A summary of our results (and an outline for this paper) is as follows.
\begin{itemize}
	\item In Section~\ref{sec:ub_symmetric_set_difference}, we show that under certain conditions the sample normalized cut $\Phi_{n,r}(\mc{C}[X])$ and conductance $\Psi_{n,r}(\mc{C}[X])$ are close to their population level counterparts $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$. As a consequence, we find that $\Delta(\wh{C},\mc{C}[X])$ is small whenever $\Phi_{\Pbb,r}(\mc{C})$ is small relative to $\Psi_{\Pbb,r}(\mc{C})^2$, giving the first population-level guarantees for local clustering in a nonparametric statistical context.
\end{itemize}
In Section~\ref{sec:ppr_density_cluster}, we focus on the special case where the candidate cluster $\mc{C} = \mc{C}_{\lambda}$ is a \emph{$\lambda$-density cluster}---that is, a connected component of the upper level set $\{x: f(x) \geq \lambda\}$---an object of central interest in the statistical clustering literature, dating back to the work of \citet{hartigan1981}.
\begin{itemize}
	\item In Section~\ref{subsec:recovery_well-conditioned_density_clusters}, we derive specific upper bounds on the normalized cut $\Phi_{\Pbb,r}(\mc{C}_{\lambda})$ and conductance $\Psi_{\Pbb,r}(\mc{C}_{\lambda})$. These upper bounds depend on $\lambda$ as well as some other natural parameters, and immediately result in an upper bound on $\Delta(\wh{C},\mc{C}[X])$ in the density cluster setting (Theorem~\ref{thm:density_cluster_volume_ssd_ub}).
	\item In Section~\ref{sec:lower_bound}, we prove a negative result: we give a hard distribution $\Pbb$ with corresponding density cluster $\mc{C}_{\lambda}$ for which the symmetric set difference $\Delta(\wh{C},\mc{C}_{\lambda}[X])$ is provably large.
\end{itemize}
In Section~\ref{sec:experiments} we empirically investigate some of our conclusions, before ending with some discussion in Section~\ref{sec:discussion}.

\subsection{Related Work}
\label{subsec:related_work}
We conclude this section by summarizing some related work (in addition to the background already given above).

\paragraph{Spectral clustering.}In the stochastic block model (SBM), arguably one of the simplest models of network formation, edges between nodes independently occur with probability based on a latent community membership. In the SBM, the ability of spectral algorithms to perform clustering---or community detection---is well-understood, dating 
back to \citet{mcsherry2001} who gives conditions under which the entire
community structure can be recovered. In more recent work, \citet{rohe2011}
upper bound the fraction of nodes misclassified by a spectral algorithm for the
high-dimensional (large number of blocks) SBM, and \citet{lei2015} extend these
results to the sparse (low average degree) regime. Relatedly,
\citet{clauset08,balakrishnan2011,li2018}, analyze the misclassification rate
when the block model exhibits some hierarchical structure. The framework we
consider, in which nodes correspond to data points sampled from an underlying 
density, and edges between nodes are formed based on geometric proximity, is
quite different than the SBM, and therefore so is our analysis.

@@

\paragraph{Density clustering.} For a given threshold $\lambda \in (0,\infty)$, let \smash{$\mathbb{C}_f(\lambda)$} denote the connected components of the density upper level set $\{x \in \Rd: f(x) > \lambda\}$. In the density clustering problem, initiated by~\cite{hartigan1975}, the goal is to recover $\mathbb{C}_{f}(\lambda)$. By now, density clustering (and the related problem of level-set estimation) are relatively well-understood. \citet{polonik1995,rigollet2009} study density clustering under the symmetric set difference
metric, \citet{tsybakov1997,singh2009} describe minimax optimal level-set
estimators under Hausdorff loss and
\citet{hartigan1981,chaudhuri2010,balakrishnan2013,kpotufe11} consider
consistent estimation of the cluster tree $\{\mathbb{C}_f(\lambda): \lambda \in (0,\infty)\}$. \textcolor{red}{Add, at a minimum, references to Larry + Ale and Ingo's papers.}

We emphasize that our goal is
not to improve on these results, nor to offer a better algorithm for level set
estimation, but rather to better understand the implications of our general theory by applying it within an already well-studied framework. Indeed, seen as a density clustering algorithm, PPR has none 
of the optimality guarantees found in the aforementioned works. We should also note that since we study a local algorithm, our interest will be in a local version of the density clustering problem, where the goal is to recover a single density cluster $\mc{C}_{\lambda} \in \mathbb{C}_f(\lambda)$.

\section{Recovery of a generic cluster with PPR}
\label{sec:ub_symmetric_set_difference}

In the main result (Theorem~\ref{thm:volume_ssd_ub}) of this section, we give a high probability upper bound on $\Delta(\wh{C}, \mc{C}[X])$, in terms of the population normalized cut $\Phi_{\Pbb,r}(\mc{C})$ and conductance $\Psi_{\Pbb,r}(\mc{C})$. We build to this theorem slowly, giving new structural results in two distinct directions. First, we build on some previous work (mentioned in the introduction) to relate $\Delta(\wh{C}, \mc{C}[X])$ to the sample normalized cut $\Phi_{n,r}(\mc{C}[X])$ and conductance $\Psi_{n,r}(\mc{C}[X])$. Second, we argue that when $n$ is large, $\Phi_{n,r}(\mc{C}[X])$ and $\Psi_{n,r}(\mc{C}[X])$ can be suitably bounded by their population-level analogues $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$.

\subsection{PPR cluster recovery: the fixed graph case}
\label{subsec:ppr_cluster_recovery_fixed_graph}
When PPR is run on a fixed graph $G = (V,E)$ with the goal of recovering a candidate cluster $C \subset V$, \cite{zhu2013} provide the sharpest known bounds on the volume of the symmetric set difference between the cluster estimate $\wh{C}$ and candidate cluster $C$. Since these results will play a major part in our analysis, in Lemma~\ref{lem:zhu} we restate them for the convenience of the reader.\footnote{Lemma~\ref{lem:zhu} improves on Lemma~3.4 of~\cite{zhu2013} by some constant factors, and for completeness we prove Lemma~\ref{lem:zhu} in our appendix. Nevertheless, to be clear the essential idea of Lemma~\ref{lem:zhu} is no different than that of~\cite{zhu2013}, and we do not claim any novelty.}

@@

\subsection{Improved bounds on mixing time} 
\label{subsec:mixing_time}
Having reviewed the conclusions of~\cite{zhu2013}, we return now to our own setting, where the data is not a fixed graph $G$ but instead random samples $\{x_1,\ldots,x_n\}$, and our goal is to recover a candidate cluster $\mc{C} \subset \Rd$. Ideally, we would like to apply~\eqref{eqn:zhu_ub2} with $C = \mc{C}[X]$ and $G = G_{n,r}$, replace $\Phi_{n,r}(\mc{C}[X])$ and $\Psi_{n,r}(\mc{C}[X])$ by $\Phi_{\Pbb,r}(\mc{C})$ and $\Psi_{\Pbb,r}(\mc{C})$ inside~\eqref{eqn:zhu_ub2}, and thereby obtain an upper bound on $\Delta(\wh{C};\mc{C}[X])$ that depends only on $\Pbb$ and $\mc{C}$. Unfortunately, however, there is a catch: when the graph $G = G_{n,r}$ and the candidate cluster $C = \mc{C}[X]$, as $n \to \infty$ the sample normalized cut $\Phi_{n,r}(\mc{C}[X])$ and conductance $\Psi_{n,r}(\mc{C}[X])$ each converge to their population-level analogues, but $\pi_{\min}\bigl(G_{n,r}\bigl[\mc{C}[X]\bigr]\bigr) \asymp 1/n$. Therefore the right hand side of~\eqref{eqn:zhu_ub2} grows at a $\log n$ rate, rendering~\eqref{eqn:zhu_ub2} a vacuous upper bound when the number of samples is sufficiently large.

@@

To establish these bounds, we impose the following regularity conditions on $\wt{\Pbb}$ and $\mc{C}$.
\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:domain} 
	The candidate cluster $\mc{C} \subseteq \Rd$ for $d \geq 2$, is a bounded, connected, open set with Lipschitz boundary. 
	\item 
	\label{asmp:bounded_density} 
	The distribution $\wt{\Pbb}$ has a density $\wt{f}: \mc{C} \to (0,\infty)$ such that there exist $f_{\min} \leq 1 \leq f_{\max}$ for which
	\begin{equation*}
	(\forall x \in \mc{C})~~ f_{\min} \leq \wt{f}(x) \leq f_{\max}.
	\end{equation*}
\end{enumerate}
In what follows, we use $b_1,b_2,\ldots$ and $B_1,B_2,\ldots$ to refer to constants that may depend on $\Pbb$ and $\mc{C}$, but do not depend on $n$ or $\delta$. We explicitly keep track of all constants in our proofs.
\begin{proposition}
	\label{prop:sample_to_population_1}
	Fix $\delta \in (0,1/3)$. Suppose $\mc{C}$ and $\wt{\Pbb}$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_density}. There exist positive constants $b_1$ and $b_2$ such that each of the following statements hold.
	\begin{itemize}
		\item With probability at least $1 - 3\exp\{-b_1\delta^2n\}$,
		\begin{equation}
		\label{eqn:sample_to_population_normalized_cut}
		\Phi_{n,r}(\mc{C}[X]) \leq (1 + 3\delta) \Phi_{\Pbb,r}(\mc{C}).
		\end{equation}
		\item With probability at least $1 - (n + 2)\exp\{-b_2\delta^2n\}$,
		\begin{equation}
		\label{eqn:sample_to_population_local_spread}
		s_{n,r}(\mc{C}[X]) \geq (1 - 3\delta) s_{\Pbb,r}(\mc{C}).
		\end{equation}
	\end{itemize}
\end{proposition}
Let $p_d := 3/4$ if $d = 2$, and else $p_d := 1/d$ for $d \geq 3$. 
\begin{proposition}
	\label{prop:sample_to_population_2}
	Fix $\delta \in (0,1/2)$. Suppose $\mc{C}$ and $\wt{\Pbb}$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_density}. Then there exist positive constants $B_1$,$B_2$, and $b_3$ such that for any $n \in \mathbb{N}$ for which
	\begin{equation}
	\label{eqn:sample_to_population_conductance_sample_complexity}
	\frac{(\log n)^{p_d}}{n^{1/d}} \leq \delta \cdot B_1,
	\end{equation}
	the following inequality holds with probability at least $1 - B_2/n - (n + 2)\exp\{-b_3n\}$:
	\begin{equation}
	\label{eqn:sample_to_population_conductance}
	\Psi_{n,r}(\mc{C}[X]) \geq (1 - 2\delta) \Psi_{\Pbb,r}(\mc{C}).
	\end{equation}
\end{proposition}

@@

\subsection{Cluster recovery}
\label{subsec:cluster_recovery}
As is typical in the local clustering literature, our algorithmic results will be stated with respect to specific ranges of each of the user-specified
parameters. In particular, for $\delta \in (0,1/4)$ and a candidate cluster $\mc{C} \in \Rd$, we require that some of the tuning parameters of Algorithm~\ref{alg:ppr} be chosen within specific ranges,
